---
title: "Seasonality"
author: "Andreas Hansen"
date: "2023-02-20"
output: html_document
---

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)
library(Kendall)
```

```{r}
prices <- read.csv("data/price_clean.csv")
con <- read.csv("data/con_clean.csv")
prod <- read.csv("data/prod_clean.csv")
```

```{r}
prices %>% 
  filter(price_area == "DK1")  %>% 
  mutate(date = lubridate::as_datetime(hour_utc)) %>% 
  mutate(value = round(as.numeric(spot_price_dkk)), digits = 0) %>% 
  select(date, value)    %>% 
  drop_na()  %>% 
  write_csv2('price_final.csv')
```

```{r}
prices <- prices %>%
  filter(price_area == "DK1") %>%
  arrange(hour_utc)

con <- con %>%
  arrange(hour_utc)

prod <- prod %>%
  filter(price_area == "DK1") %>%
  arrange(hour_utc)

splitfn <- function(x) {
  return(strsplit(strsplit(x, '-')[[1]][3], ' ')[[1]][1])
}
prices["Hour"] <- hour(prices$hour_utc)
prices["Wday"] <- wday(prices$hour_utc, label=T)
prices["Mday"] <- sapply(prices$hour_utc, splitfn)
prices["Week"] <- week(prices$hour_utc)
prices["Month"] <- month(prices$hour_utc)
prices["Year"] <- year(prices$hour_utc)
```


## Seasonality

```{r}
prices_seas <- prices

plot(prices_seas$spot_price_dkk, type='line')
```

The mean price for each hour has large variations so we might suspect some hourly seasonality.
```{r}
prices_seas <- prices %>%
  group_by(Hour) %>%
  summarize(mean = mean(spot_price_dkk), sd = sd(spot_price_dkk))

head(prices_seas, n=24)
```

The mean price for each weekday has large variation so we might suspect weekday seasonality.
```{r}
prices_seas <- prices %>%
  group_by(Wday) %>%
  summarize(mean = mean(spot_price_dkk), sd = sd(spot_price_dkk))

head(prices_seas, n=7)
```

```{r}
prices_seas <- prices %>%
  group_by(Week) %>%
  summarize(mean = mean(spot_price_dkk), sd = sd(spot_price_dkk))

head(prices_seas, n=52)
```

```{r}
prices_seas <- prices %>%
  group_by(Year) %>%
  summarize(mean = mean(spot_price_dkk), sd = sd(spot_price_dkk))

head(prices_seas, n=5)
```


```{r}
prices %>%
  filter(hour_utc >= "2020-01", hour_utc < "2020-02") %>%
  group_by(Hour) %>%
  ggplot(aes(x=date(hour_utc), 
             y=spot_price_dkk, 
             group=factor(Hour), 
             color=factor(Hour))) + 
  geom_line()
```

## Hourly and daily seasonality

The mean price for each hour grouped by week day shows that the price varies greatly through out a day indication hourly seasonality. We also see a similar price pattern through out a day for all week days.
```{r}
prices %>%
  group_by(Hour, Wday) %>%
  summarize(hour=Hour, day=Wday, week=Week, month=Month, year=Year, mean_price=mean(spot_price_dkk)) %>%
  ggplot(aes(x=hour, 
             y=mean_price, 
             group=factor(day), 
             color=factor(day))) + 
  geom_line()
```

### Weekly seasonality

The mean price for each week day varies greatly through out a year.
```{r}
prices %>%
  group_by(Wday, Week) %>%
  summarize(hour=Hour, day=Wday, week=Week, month=Month, year=Year, mean_price=mean(spot_price_dkk)) %>%
  ggplot(aes(x=week, 
             y=mean_price, 
             group=factor(day), 
             color=factor(day))) + 
  geom_line()
```


### Monthly seasonality

```{r}
prices %>%
  group_by(Wday, Month) %>%
  summarize(hour=Hour, day=Wday, week=Week, month=Month, year=Year, mean_price=mean(spot_price_dkk)) %>%
  ggplot(aes(x=day, 
             y=mean_price, 
             group=factor(Month), 
             color=factor(Month))) + 
  geom_line()
```

Similar to weeks, then the price varies for each month. The pattern is similar for each week day for each month during a year.
```{r}
prices %>%
  group_by(Wday, Month) %>%
  summarize(hour=Hour, day=Wday, week=Week, month=Month, year=Year, mean_price=mean(spot_price_dkk)) %>%
  ggplot(aes(x=month, 
             y=mean_price, 
             group=factor(day), 
             color=factor(day))) + 
  geom_line()
```

### Yearly seasonlity

```{r}
prices %>%
  group_by(Month, Year) %>%
  summarize(hour=Hour, day=Wday, week=Week, month=Month, year=Year, mean_price=mean(spot_price_dkk)) %>%
  ggplot(aes(x=month, 
             y=mean_price, 
             group=factor(year), 
             color=factor(year))) + 
  geom_line()
```


### ACF

```{r}
acf(prices$spot_price_dkk)
pacf(prices$spot_price_dkk)
```

```{r}
acf(diff(prices$spot_price_dkk, 1))
pacf(diff(prices$spot_price_dkk, 1))
```

Ikke en god metode til at checke for sæson.
```{r}
prices_seas <- prices %>%
  filter(price_area == "DK1")

acf(prices_seas$spot_price_dkk, lag.max=100)
```

Vi kan se en "top" pr. 12'te time (dermed et 24 timers gentagende mønster), og derfor seasonality på timebasis.


```{r}
prices_seas <- prices_seas %>%
  filter(Hour == 12)

acf(prices_seas$spot_price_dkk, lag.max=80)
```

Vi kan se en "top" pr. 7'ende dag, og derfor seasonality på dagsbasis.


```{r}
prices_seas <- prices_seas %>%
  filter(Wday == "Ons")

acf(prices_seas$spot_price_dkk, lag.max=50)
```

Det er sværere at se et præcist gentagende mønster nu på ugentlig basis (hver 4. observation) og på månedlig basis (hver 12. observation).


